\protect Connecting every bit of knowledge: \\ The Structure of Wikipedia's first link network

### 2016-02-20 writeup/wikipedia-network.tex
graphics and language ok

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
are related and organized.  
We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
We traverse every possible path through the network, 
measuring the accumulation of first links, path lengths, basins, cycles, and the influence a particular article exerts in shaping the 
network.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general 
culminating around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
Wikipedia's ever growing store of knowledge.

@@

TODO: 
\begin{itemize}
    \item add the what to scale-free distribution in the abstract
    \item fix end of second paragraph in Results Section\\
    * power-law biz: rank exponent (slope of log-log), denoted $\alpha$ \\
    * size exponent corresponds to the CDF (how many people have an income greater than x)
    = $\gamma = \frac{1}{\alpha} + 1$.
\item fix exponent reporting (with alpha) \\
    * figure out what gamma size freq means
\end{itemize}

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution.}
We fit the in-degree distribution to a linear $\log_{10}-\log_{10}$ model. The result is a 
an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
}
  \label{fig:degree distribution}
\end{figure}

The FLN's in-degree exhibits a decaying power law distribution where a few articles 
receive most direct first references, while most articles receive few or none.
The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
When fit with a linear model on a log-log scale (log(rank) versus log(in-degree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.\\

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal visits. 
The model explains $86\%$ of the variation in the data, yielding an r-value of $-0.93$ 
and a power law exponent of $-0.64$. The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).}
  \label{fig:Distribution of Visits}

@@

In Fig.~\ref{fig:Distribution of Visits}, we show a log-log plot for the entire dataset:traversal visits against rank. 
We observe a strong linear fit, with a Pearson's Correlation Coefficient of -.93 and 
a power-law exponent of $-0.636$. A handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
We find two regimes with the top regime (log(rank) < 4) 
well-explained by a linear fit, yielding an r-value of $-0.99$ and a 
power-law exponent of $-1.08$. The top regime 
explains the distribution of traversal funnels for the $17, 821$ 
articles with at least one traversal funnel. The bottom 
horizontal regime corresponds to the more than $99\%$ of articles
which hold zero traversal funnels.}
  \label{fig:Funnels Distribution}
\end{figure}
Ranking articles by the number of traversal funnels we find 
``Philosophy'' to be by far the highest-ranking article with 
$7.37$ million paths
(see Fig.~\ref{fig:Funnels}).
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by more than two orders of magnitude.
The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
and similar articles accumulates the overwhelming proportion of its 
references through ``Philosophy'': $7.37$ million of the $7.4$ million references
are funneled through ``Philosophy''.
Second on the list of highest-ranking articles by traversal funnels is 
``Presentation'' with only $30$ thousand paths. Similarly abstract 
concepts also rank highly such as ``Tree of life'' (30 thousand), 
``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).

@@

As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
an more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When fit to a log-log linear model we find the $99\%$ of articles with zero
traversal funnels form one regime (with log(rank) less than 4).
The top regime, corresponding to the $17, 821$ articles with at least one 
traversal funnel strongly fits a linear model with an r-value of $-0.99$. 
The resulting power-law exponent is $-1.08$. Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

>>>>

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in an article, we connect entries to form a directed network: Wikipedia's First Link Network. 
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
are related and organized.  
We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
We traverse every possible path through the network, 
measuring the accumulation of first links, path lengths, basins, cycles, and the influence a particular article exerts in shaping the 
network.
We discover many scale-free distributions describing path length, 
accumulation, and influence; find Philosophy at a salient center; and uncover a flow from specific to general 
culminating around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

@@

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution.}
We fit the in-degree distribution with a linear $\log_{10}-\log_{10}$ model. 
The result is a 
an excellent fit with a Pearson's Correlation Coefficient of $-0.98$, yielding a 
power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$. 
}
  \label{fig:degree distribution}
\end{figure}

The FLN's in-degree exhibits a decaying power law distribution where a few articles 
receive most direct first references, while most articles receive few or none.
The average in-degree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
Less than $1\%$ of articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
We fit the in-degree distribution against each article's rank with a log-log (base 10) linear regression model.
We obtain a Pearson's Correlation Coefficient of $-0.98$
with a corresponding  power law rank exponent $\alpha = -0.788$ and a size exponent $\gamma = -0.266$.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits.}
We fit the distribution of traversal visits against article rank with a log-log (base 10) linear regression model log-log model and uncover two regimes: 
with a cutoff rank of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
The horizontal flattening around the highest
ranking articles is a result of the cyclic structure (see discussion on cycles).}
  \label{fig:Distribution of Visits}

@@

In Fig.~\ref{fig:Distribution of Visits}, we show a log-log (base 10) plot for the entire dataset: traversal visits against rank. 
We fit the data to a two regime log-log linear regression with a cutoff rank 
of $10^{5}$.
The top regime ($\log_{10}$(rank) $< 5$) has a Pearson's Correlation 
Coefficient of $-0.99$ with a corresponding power law rank exponent 
$\alpha = -1.23$ and a size exponent $\gamma = 0.187$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/funnels_distribution.png}
  \caption{
    \textbf{Distribution of Traversal Funnels.}
We fit the distribution with a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal funnels. 
We find two regimes with the top regime (log(rank) < 4) 
well-explained by a linear fit, yielding Pearson's 
Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$. The top regime 
explains the distribution of traversal funnels for the $17, 821$ 
articles with at least one traversal funnel. The bottom 
horizontal regime corresponds to the more than $99\%$ of articles
which hold zero traversal funnels.}
  \label{fig:Funnels Distribution}
\end{figure}
Ranking articles by the number of traversal funnels we find 
``Philosophy'' to be by far the highest-ranking article with 
$7.37$ million paths
(see Fig.~\ref{fig:Funnels}).
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by more than two orders of magnitude.
The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
and similar articles accumulates the overwhelming proportion of its 
references through ``Philosophy'': $7.37$ million of the $7.4$ million references
are funneled through ``Philosophy''.
Second on the list of highest-ranking articles by traversal funnels is 
``Presentation'' with only $30$ thousand paths. Similarly abstract 
concepts also rank highly such as ``Tree of life'' (30 thousand), 
``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).

@@

As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
an more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When fit to a log-log linear model we find the $99\%$ of articles with zero
traversal funnels form one regime (with log(rank) less than 4).
The top regime, corresponding to the $17, 821$ articles with at least one 
traversal funnel strongly fits a linear model with
a Pearson's Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$. Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.
